# Preface {.unnumbered}


```{r}
# this is just a test, delete me
knitr::read_chunk(path = "R/data_app.R")
```

<!--# about the book -->

```{r}
#| include: false
# library(spelling)
# library(devtools)
```

## Who is this for

This tutorial is aimed at data scientists / analysts who want to write their first `teal` application. Some R knowledge is required although no exposure to `shiny` is expected.

The `teal` package aims to make the creation of (teal-specific) shiny applications intuitive --- without the need to understand how shiny works exactly. Shiny knowledge is useful once you want to either modify an application, or write your own teal module from scratch.

## How is it structured

All chapters in this book follow a common structure: First, the learning goals of a chapter is defined. We then go through the step-by-step process of using specific `teal` modules. In each chapter we combine what we learned into one code chunk with which you can create a working teal application (*toy app*). As every `teal` app requires data and meta-data, the *toy app* created in the introduction will be the fundamental building block of the apps in the subsequent chapters. In the *toy apps* created in each chapter you should be able to understand every line of code. The code for a more developed app (but built for the same purpose) and a deployed instance of that app is linked at the end of each chapter. What you learned in this chapter should give you a good understanding of how the more developed application works.

## What will you learn

You will learn the structure of `teal` applications and the structure of the corresponding code.

After this tutorial you should be able to create your own application for your data based on existing `teal` modules. This requires an awareness of --- some --- available modules, where to find them, their intent, how to use them, and where to find the rest. Further, you should be able to customize existing modules to suit your specific needs.

## Where can you find more

-   link to other manuals for teal

<!--# This is a collection of materials used for the introductory content of teal found on the [pharmaverse youtube channel](https://www.youtube.com/@pharmaverse6572). -->
